### 2024-06-14 products/manuscript/Manuscript.qmd

---
title: "Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset "
subtitle: ""
author: Antonio Flores
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
library(here)
library(knitr)
```

# Summary/Abstract
This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.

{{< pagebreak >}}

# Introduction

## General Background Information
Providing businesses with a model that will allow prioritization of consumers could save time and allow for a more accurate use of resources.

## Description of data and data source
The data was donated on 2/13/2012. It was collected from phone call marketing campaigns performed by a Portuguese banking institution.I have accessed this data from the UC Irvine Machine Learning Repository.

There are 45,212 records, and includes: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history. Additionally, the classification variable is whether or not the person subscribed to a term deposit. There are 17 features in total.

Among the variables are a handful of features relating to the marketing campaign itself. For example, included are the day, month, and duration of the call, the number of contacts performed during the campaign (campaign), the number of days since the client was last contacted (pdays), the number of contacts performed before this campaign (previous), and finally the outcome of precious marketing campaigns (poutcome).

## Questions/Hypotheses to be addressed

The research question I plan to address with my analysis is which features or combination of features are the best predictors of consumers making a deposit. The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns. I don't currently have specific predictors in mind to begin with, as I plan to test all of the available features while also looking at some combinations or concatenation of features.

{{< pagebreak >}}

# Methods

I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already. Then I will test for correlation between different features and proceed from there.

## Schematic of workflow

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
```

## Data aquisition
Data retrieved from UCI ML Repository

## Data import and cleaning

### Reading in the Data

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Data Snapshot"
#| echo: FALSE
data_location <- here::here("data","raw-data","bank-full.csv")
raw=read.csv(data_location, header = TRUE, sep=";")
head(raw)
```

### Dimensions:
dim(raw)

```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Data Dimensions"
#| echo: FALSE
dim(raw)
```

### Describing raw data
str(raw)

```{r}
#| label: tbl-summarytable3
#| tbl-cap: "Data Description"
#| echo: FALSE
str(raw)
```

## Statistical analysis

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

@fig-result shows a boxplot figure comparing balance levels across job types.

```{r}
#| label: fig-result
#| fig-cap: "Job type and bank account balance"
#| echo: FALSE
knitr::include_graphics(here("results","figures","job-balance-scatter.png"))
```

@fig-result1 shows a bar chart showing education levels of the data.

```{r}
#| label: fig-result1
#| fig-cap: "Education level"
#| echo: FALSE
knitr::include_graphics(here("results","figures","education-barchart.png"))
```

@fig-result2 shows a scatter plot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Age and bank account balance stratified by marital status"
#| echo: FALSE
knitr::include_graphics(here("results","figures","age-balance-stratified.png"))
```

## Basic statistical analysis

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
#knitr::kable(resulttable2)
```

## Full analysis

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

{{< pagebreak >}}

# References