### 2024-06-22 products/manuscript/Manuscript.qmd

---
title: "Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset "
subtitle: ""
author: Antonio Flores
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
library(here)
library(knitr)
```

# Summary/Abstract
This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank. After the data has been cleaned and prepared for analysis, exploratory data analysis will be performed to gather more information regarding the shape, size, and behavior of different variables, to gauge their usefulness in a prediction model. Additionally, several different statistical tests will be implemented to identify which classes within different variables affect the outcome variable. Finally (_haven't started on this yet_), with some additional data preprocessing, this project will test different machine learning models and conclude with the resulting findings.

{{< pagebreak >}}

# Introduction

## General Background Information
Providing businesses with a model that will allow prioritization of consumers and/or demographics has great potential in improving resource management and future marketing campaigns, as well as increasing efficient spending.

## Description of data and data source
The data was donated on 2/13/2012. It was collected from phone call marketing campaigns performed by a Portuguese banking institution.I have accessed this data from the UC Irvine Machine Learning Repository.

There are 45,212 records, and includes: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history. Additionally, the classification variable is whether or not the person subscribed to a term deposit. There are 17 features in total.

Among the variables are a handful of features relating to the marketing campaign itself. For example, included are the day, month, and duration of the call, the number of contacts performed during the campaign (campaign), the number of days since the client was last contacted (pdays), the number of contacts performed before this campaign (previous), and finally the outcome of precious marketing campaigns (poutcome).

## Questions/Hypotheses to be addressed

The research question I plan to address with my analysis is: which features or combination of features are the best predictors of consumers making a deposit? The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns. Currently, I plan to investigate all variables, but I am specifically interested in both job type, education and age.

{{< pagebreak >}}

# Methods

Cleaning -> This includes converting numeric variables to factors, changing column names, and converting the positive outcome varible to 1 and 0.

Exploring -> Using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.

## Data aquisition
The dataset for this project was retrieved from UCI ML Repository in CSV form.
Additionally, I created a codebook based on data from the same source.

## Data import and cleaning

### Reading in the Data

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Data Snapshot"
#| echo: FALSE
data_location <- here::here("data","raw-data","bank-full.csv")
raw=read.csv(data_location, header = TRUE, sep=";")
head(raw)
```

### Dimensions:
dim(raw)

```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Data Dimensions"
#| echo: FALSE
dim(raw)
```

### Describing raw data
str(raw)

```{r}
#| label: tbl-summarytable3
#| tbl-cap: "Data Description"
#| echo: FALSE
str(raw)
```

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

@fig-result shows a boxplot figure comparing balance levels across job types.

```{r}
#| label: fig-result
#| fig-cap: "Job type and bank account balance"
#| echo: FALSE
knitr::include_graphics(here("results","figures","job-balance-scatter.png"))
```

@fig-result1 shows a bar chart showing education levels of the data.

```{r}
#| label: fig-result1
#| fig-cap: "Education level"
#| echo: FALSE
knitr::include_graphics(here("results","figures","education-barchart.png"))
```

@fig-result2 shows a scatter plot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Age and bank account balance stratified by marital status"
#| echo: FALSE
knitr::include_graphics(here("results","figures","age-balance-stratified.png"))
```

@fig-result4 shows barplot of age data for the full dataset

```{r}
#| label: fig-result4
#| fig-cap: "Age data for those with positive classifcation"
#| echo: FALSE
knitr::include_graphics(here("results","figures","yes-no-age.png"))
```

@fig-result3 shows barplot of age data for only those who subscribed to a term deposit

```{r}
#| label: fig-result3
#| fig-cap: "Age data for those with positive classifcation"
#| echo: FALSE
knitr::include_graphics(here("results","figures","yes-age.png"))
```

@fig-result5 shows barplot of most common days of the month to record a positive outcome.

```{r}
#| label: fig-result5
#| fig-cap: "Days of the month for positive outcomes"
#| echo: FALSE
knitr::include_graphics(here("results","figures","days-yes.png"))
```

@fig-result6 Job type for count of positive outcomes.

```{r}
#| label: fig-result6
#| fig-cap: "Job type for count of positive outcomes."
#| echo: FALSE
knitr::include_graphics(here("results","figures","jobs-count-yes.png"))
```

@fig-result7 Job type for percent of positive outcomes per job type.

```{r}
#| label: fig-result7
#| fig-cap: "Job type for percent of positive outcomes per job type."
#| echo: FALSE
knitr::include_graphics(here("results","figures","jobs-percent-yes.png"))
```

## Basic statistical analysis

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
#knitr::kable(resulttable2)
```

#WORK IN PROGRESS

## Full analysis

_Pending Data Preprocessing and ML Modeling_
_I plan to use recipies and some of the tinymodels for this section_

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

{{< pagebreak >}}

# References