### 2024-07-13 code/processing-code/processingfile-v1 - my version.qmd

---
title: "Preparing data for analysis"
author: "Antonio Flores"
date: "2024-13-06"
output: html_document
---

# Setup

The following packages are needed for this data processing

```{r}
#| echo: false
#| output: false
library(readxl) 
library(dplyr) 
library(tidyr) 
library(skimr) 
library(here) 
```

# Data loading

```{r}
data_location <- here::here("data","raw-data","bank-full.csv")
rawdata <- read.csv(data_location, header = TRUE, sep=";")
```

# Check data

Examining metadata by accessing the codebook. This codebook was taken directly from the data repository housing the chosen dataset. The decision was made to keep this in a separate xlsx file from the dataset so that the dataset could remain in the identical form in which it was sourced (CSV format).

```{r}
codebook_location <- 
here::here("data","raw-data","codebook.xlsx")

codebook <- readxl::read_excel(codebook_location, sheet ="Codebook")
print(codebook)
```

Various ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning the data

We are mainly converting character variables into factors so we can do some categorical analysis. I had initially included converting months to a numeric variable but realized that it was fine as a factor variable.

```{r}
d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(y = ifelse(y == "yes", 1, 0)) |>
  mutate(y = as.factor(y)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(day = as.factor(day)) |>
  mutate(y = as.factor(y)) %>% 
  select(!duration)

colnames(d1)[16] <- "y_termSubscribed"

summary(d1)
```

All done, data is clean now.

Assigning data to output variable.

```{r}
processeddata1 <- d1
```

# Save data

```{r}
save_data_location <- here::here("data","processed-data","processeddata1.rds")
saveRDS(processeddata1, file = save_data_location)
```

# Attempting to convert categorical variables to dummy

```{r}
library(caret)

testing_dummy = processeddata1 %>% 
  select(job, marital, education, default, housing, loan, contact) %>% 
  mutate(ID = row_number())

dummy <- dummyVars("~.", data=testing_dummy)
df1 <- predict(dummy, newdata = testing_dummy)

df1 = as.data.frame(df1)

numeric_vars = processeddata1 %>% 
  select(age, balance, day, y_termSubscribed) %>% 
  mutate(ID = row_number())

d2 <- merge(df1,numeric_vars, by = "ID")

```

```{r}
save_data_location <- here::here("data","processed-data","processeddata2.rds")
saveRDS(d2, file = save_data_location)
```

# Attempting to convert categorical variables to numeric factors

```{r}
library(caret)

d3 = processeddata1 |>
  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
  mutate(job = as.factor(unclass(job))) |>
  mutate(marital = as.factor(unclass(marital))) |>
  mutate(education = as.factor(unclass(education))) |>
  mutate(default = as.factor(unclass(default))) |>
  mutate(housing = as.factor(unclass(housing))) |>
  mutate(loan = as.factor(unclass(loan)))

summary(d3$education)
```

```{r}
save_data_location <- here::here("data","processed-data","processeddata3.rds")
saveRDS(d3, file = save_data_location)
```

# Attempting to convert categorical variables to numeric

```{r}
library(caret)

d4 = processeddata1 |>
  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
  mutate(job = unclass(job)) |>
  mutate(marital = unclass(marital)) |>
  mutate(education = unclass(education)) |>
  mutate(default = unclass(default)) |>
  mutate(housing = unclass(housing)) |>
  mutate(loan = unclass(loan))|>
  mutate(day = unclass(day))

```

```{r}
save_data_location <- here::here("data","processed-data","processeddata4.rds")
saveRDS(d4, file = save_data_location)
```

# Attempting to convert categorical variables to true numeric

```{r}
library(caret)

d5 = processeddata1 |>
  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
  mutate(job = as.numeric(job)) |>
  mutate(marital = as.numeric(marital)) |>
  mutate(education = as.numeric(education)) |>
  mutate(default = as.numeric(default)) |>
  mutate(housing = as.numeric(housing)) |>
  mutate(loan = as.numeric(loan))|>
  mutate(day = as.numeric(day))

```

```{r}
save_data_location <- here::here("data","processed-data","processeddata5.rds")
saveRDS(d5, file = save_data_location)
```